---
title: "Global Baseline Estimate"
author: "Nwe Oo Mon (Nina)"
date: "2024-09-16"
output: html_document
---

# Creating MySQL database
#### Separate tables namely critics, movies, and movie_ratings were created and the data was loaded from csv files. The sql codes can be check through the link [Global_baseline_estimate](https://github.com/Nweoomon/Global_baseline_estimate/blob/main/movies_rating_2.sql).

# Loading the information from the SQL database into an R dataframe

#### Installing and Loading Required Packages
```{r setup}
if (!requireNamespace("DBI", quietly = TRUE)) install.packages("DBI")
if (!requireNamespace("RMariaDB", quietly = TRUE)) install.packages("RMariaDB")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")

library(DBI)
library(RMariaDB)
library(dplyr)

```

#### Connecting to the MySQL Database and Loading Data

```{r connect-to-mysql, echo=TRUE, message=FALSE, warning=FALSE}

con <- dbConnect(
  RMariaDB::MariaDB(),
  user = 'root',               
  password = Sys.getenv("MYSQL_PWD"),    
  dbname = 'movies_rating2',     
  host = 'localhost',           
  port = 3306                   
)

# Check if the connection is successful
if (!is.null(con)) {
  print("Connected successfully!")
} else {
  print("Connection failed.")
}

# Execute the SQL query and fetch the results into an R data frame
query <- "SELECT Critics.CriticName, Movies.MovieName, Ratings.Rating
          FROM Ratings
          JOIN Critics ON Ratings.CriticID = Critics.CriticID
          JOIN Movies ON Ratings.MovieID = Movies.MovieID;"

# Fetch the data
result <- dbGetQuery(con, query)

# Display the results
glimpse (result)

# Close the database connection
dbDisconnect(con)
```
#### Global Baseline Estimate Calculation

```{r GBE}

# Calculating the mean movie rating
mean_movie_rating <- mean(result$Rating, na.rm = TRUE)

# Calculate movie average ratings
movie_avg <- result %>%
  group_by(MovieName) %>%
  summarize(movie_avg = mean(Rating, na.rm = TRUE)) %>%
  ungroup()

# Calculate user average ratings
user_avg <- result %>%
  group_by(CriticName) %>%
  summarize(user_avg = mean(Rating, na.rm = TRUE)) %>%
  ungroup()

# Calculate movie rating relative to average
movie_avg_relative <- movie_avg %>%
  mutate(movie_avg_relative = movie_avg - mean_movie_rating)

# Calculate user rating relative to average
user_avg_relative <- user_avg %>%
  mutate(user_avg_relative = user_avg - mean_movie_rating)

# Merge these data back into the original data frame
result <- result %>%
  left_join(movie_avg_relative, by = "MovieName") %>%
  left_join(user_avg_relative, by = "CriticName")

# Ensure that all necessary columns are numeric
result <- result %>%
  mutate(
    movie_avg = as.numeric(movie_avg),
    user_avg = as.numeric(user_avg),
    movie_avg_relative = as.numeric(movie_avg - mean_movie_rating),
    user_avg_relative = as.numeric(user_avg - mean_movie_rating)
  )

# Calculate the global baseline estimate for each combination
result <- result %>%
  mutate(Global_Baseline_Estimate = mean_movie_rating + movie_avg_relative + user_avg_relative)

glimpse (result)

# Filter to find Param's predicted rating for Pitch Perfect 2
param_pitch_perfect2 <- result %>%
  filter(CriticName == "Param\r" & MovieName == "PitchPerfect2\r")

# Print the result for Param's predicted rating for Pitch Perfect 2
print(param_pitch_perfect2)

```





